---
title: "Lecture 7 Introduction to Functions EDS 221"
author: "Dalila Lara"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# Lecture #7

### For loops example wtih paste0() function

```{r}
dosage <- c(10, 100, 1000)
dose_index <- vector(mode = "character", length = length(dosage))

for (i in seq_along(dosage)) {
  k <- paste0("zinc_", dosage[i])
  dose_index[i] <- k
}
```

### A nested for loop example

```{r}
# Stored vectors 
file_prefix <- c("temp", "ph", "salinity")
file_suffix <- c(1, 2, 3, 4, 5)

#nested for loop seq_along
#note i = 1, j = 2 output temp_2 etc.
for (i in seq_along(file_prefix)) {
  for(j in seq_along(file_suffix)) {
    print(paste0(file_prefix[i], "_", file_suffix[j]))
  }
}
```

### Nested loop for example with operators (\*/-+)

Interloops need to evaluated first for outerloops to be evaluated

```{r}
odds <- c(1, 3, 5)
evens <- c(2, 4, 6, 8)

for(i in seq_along(odds)) {
  for(j in seq_along(evens)) {
    print(odds[i] * evens[i])
  }
}
```

## Functions: self contained bits of code that accept user inputs(s), do something with it, and return a useful output.

Consider writing a function when you copy and paste a code block more than twice. 1. give function name to make code readable 2. requirements change update code in that one place instead of many 3. eliminate chance of mistakes with copy and paste

### Anatomy of function -R

Function name, argument values, body, and output

```{r}
#function name               #argument values 
function_name <- function(bird, dog) {
pets = bird + dog                         #body
return(pets)                              #outputs
}
```

Function Example

```{r}
exclaim_age <- function(age) {
  print(paste("I am", age, "years old!"))
}
exclaim_age(age = 10)
```

Naming functions be clear rather than short.

# Computational Session Day 4 EDS 221

#Loop Example Adding the first element to the next and so forth.

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)
for (i in seq_along(quarter_splits)) {
  half_time <- quarter_splits[i] + quarter_splits[i +1]
  print(half_time)
}
```

# Functions with conditions

```{r}
animal_age <- function(animal, age) {
  if(animal == "dog") {
    print (age * 7)
  } else if (animal == "goat") {
    print(age * 4.7)
  }
}

# Now using function with specified elements 
animal_age(animal = "goat", age = 7)
animal_age(animal = "dog", age = 2)
```

# Intro to functions

Make a function to double the user input
```{r}
double_this <- function(my_value) {
  print(2 * my_value)
}

double_this(my_value = 5)
double_this(200)
double_this(c(1, 2, 10, 100, 175))
```

Write a function that asks for users NAME and AGE and returns and exclamation like, I'm (NAME), and I am (AGE) years old.
```{r}
exclaim_ages <- function(your_name, your_age) {
  print(paste("My name is", your_name, "and I am", your_age, "years old!"))
}

exclaim_ages(your_name = "Teddy", your_age = 4)
```

Calculating animal age based on human age function
### Function and conditional statement
```{r}
convert_animal_age <- function(animal, human_age) {
  if (animal == "dog") {
    print(human_age * 7)
  } else if(animal == "goat") {
    print(human_age * 4.7)
  } else {
    print("Please enter dog or goat for animal.")
  }
}

convert_animal_age(animal = "dog", human_age = 12)
convert_animal_age(animal = "goat", human_age = 5)
convert_animal_age(animal = "pika", human_age = 2)
```

Pull information from data frame 
```{r}
#load tribble() of tidyverse
# Make a data frame with tribble() function 
dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)

#just given a dog name pull information from data frame 
#created subset data fram
#pulled specified data from subset 
dog_menu <- function(name) {
  my_subset <- filter(dog_choice, dog_name == name)
  #print(my_subset)
  print(paste("My name is", my_subset$dog_name, "and my favorite foood is", my_subset$food))
}

dog_menu(name = "Teddy")
```

### Functions meet for loops 
Function and forloops example
```{r}
#apply function essentially
means_over_columns <- function(df) {
  for (i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    column_name <- colnames(df[i])
    print(paste("The mean value of", column_name, "is", col_mean))
  }
}

means_over_columns(df = mtcars)
means_over_columns(df = faithful)
```

