---
title: "Lecture 7 Introduction to Functions EDS 221 & Lecture 8"
author: "Dalila Lara"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# Lecture #7

### For loops example wtih paste0() function

```{r}
dosage <- c(10, 100, 1000)
dose_index <- vector(mode = "character", length = length(dosage))

for (i in seq_along(dosage)) {
  k <- paste0("zinc_", dosage[i])
  dose_index[i] <- k
}
```

### A nested for loop example

```{r}
# Stored vectors 
file_prefix <- c("temp", "ph", "salinity")
file_suffix <- c(1, 2, 3, 4, 5)

#nested for loop seq_along
#note i = 1, j = 2 output temp_2 etc.
for (i in seq_along(file_prefix)) {
  for(j in seq_along(file_suffix)) {
    print(paste0(file_prefix[i], "_", file_suffix[j]))
  }
}
```

### Nested loop for example with operators (\*/-+)

Interloops need to evaluated first for outerloops to be evaluated

```{r}
odds <- c(1, 3, 5)
evens <- c(2, 4, 6, 8)

for(i in seq_along(odds)) {
  for(j in seq_along(evens)) {
    print(odds[i] * evens[i])
  }
}
```

## Functions: self contained bits of code that accept user inputs(s), do something with it, and return a useful output.

Consider writing a function when you copy and paste a code block more than twice. 1. give function name to make code readable 2. requirements change update code in that one place instead of many 3. eliminate chance of mistakes with copy and paste

### Anatomy of function -R

Function name, argument values, body, and output

```{r}
#function name               #argument values 
function_name <- function(bird, dog) {
pets = bird + dog                         #body
return(pets)                              #outputs
}
```

Function Example

```{r}
exclaim_age <- function(age) {
  print(paste("I am", age, "years old!"))
}
exclaim_age(age = 10)
```

Naming functions be clear rather than short.

# Computational Session Day 4 EDS 221

#Loop Example Adding the first element to the next and so forth.

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)
for (i in seq_along(quarter_splits)) {
  half_time <- quarter_splits[i] + quarter_splits[i +1]
  print(half_time)
}
```

# Functions with conditions

```{r}
animal_age <- function(animal, age) {
  if(animal == "dog") {
    print (age * 7)
  } else if (animal == "goat") {
    print(age * 4.7)
  }
}

# Now using function with specified elements 
animal_age(animal = "goat", age = 7)
animal_age(animal = "dog", age = 2)
```

# Intro to functions

Make a function to double the user input
```{r}
double_this <- function(my_value) {
  print(2 * my_value)
}

double_this(my_value = 5)
double_this(200)
double_this(c(1, 2, 10, 100, 175))
```

Write a function that asks for users NAME and AGE and returns and exclamation like, I'm (NAME), and I am (AGE) years old.
```{r}
exclaim_ages <- function(your_name, your_age) {
  print(paste("My name is", your_name, "and I am", your_age, "years old!"))
}

exclaim_ages(your_name = "Teddy", your_age = 4)
```

Calculating animal age based on human age function
### Function and conditional statement
```{r}
convert_animal_age <- function(animal, human_age) {
  if (animal == "dog") {
    print(human_age * 7)
  } else if(animal == "goat") {
    print(human_age * 4.7)
  } else {
    print("Please enter dog or goat for animal.")
  }
}

convert_animal_age(animal = "dog", human_age = 12)
convert_animal_age(animal = "goat", human_age = 5)
convert_animal_age(animal = "pika", human_age = 2)
```

Pull information from data frame 
```{r}
#load tribble() of tidyverse
# Make a data frame with tribble() function 
dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)

#just given a dog name pull information from data frame 
#created subset data fram
#pulled specified data from subset 
dog_menu <- function(name) {
  my_subset <- filter(dog_choice, dog_name == name)
  #print(my_subset)
  print(paste("My name is", my_subset$dog_name, "and my favorite foood is", my_subset$food))
}

dog_menu(name = "Teddy")
```

### Functions meet for loops 
Function and forloops example
```{r}
#apply function essentially
means_over_columns <- function(df) {
  for (i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    column_name <- colnames(df[i])
    print(paste("The mean value of", column_name, "is", col_mean))
  }
}

means_over_columns(df = mtcars)
means_over_columns(df = faithful)
```

# Computational Part 2 EDS 221 Day 4
#Comprehension check
Matrix created with `for loop`
```{r}
#unequal sizes in vectors then they line up, but need to be carefull if they are exactly the same.
apples <- c(1,2,3,4)
price <- c(7,10,25)
#creating empty matrix
fruit_out <- matrix(nrow = length(price), ncol = length(apples))

#how you store it tells you how it will populate
#rows populate first and then columns in this case 
for(i in seq_along(apples)) {
  for(j in seq_along(price)) {
    total_cost <- price[j] * apples[i]
    fruit_out[j, i] <- total_cost
  }
}
```

### Example Nested for loop example AKA making a matrix
```{r}
coffee <- c(1,2,3,4,5)
price <- c(3.50,4.50,5.00)
#creating empty matrix
coffee_price <- matrix(nrow = length(price), ncol = length(coffee))

#how you store it tells you how it will populate
#rows populate first and then columns in this case 
for(i in seq_along(coffee)) {
  for(j in seq_along(price)) {
    total_cost <- price[j] * coffee[i]
    coffee_price[j, i] <- total_cost
  }
}
```

Example #2 Nested for loop
```{r}
clothing <- c("shirt", "pants", "shoes")
color <- c("purple", "red", "blue", "white")

our_clothes <- matrix(nrow = length(color), ncol = length(clothing))

for(i in seq_along(clothing)) {
  for(j in seq_along(color)) {
    clothing_details <- paste(color[j], clothing[i])
    our_clothes[j,i] <-clothing_details
  }
}
#view matrix created
our_clothes
```

# Lecture 8: Functions continued - planning, building, adding complexity, useful warning, and error messages
 Developing functions
 - have a goal (draw it out)
 - make a plan and map it
 - build bit by bit (mostly when more complex)
 - check intermediates

Logistic Growth
- Models population growth as it nears a carrying capacity
Loop over different growth rates. Matrix of population size outputs look like for different times and growth rates.

Plan and pseudo code think and write code. 
```{r}
# Make function for logistic growth
#logistic_growth <- function(N0, K, r, time) {
 # Nt <- K / (1 + ((K - N0) / N0) * exp(-r * time))
  #print(Nt)
#}

# now I can create a for loop to iterate this over several population growth rates

#for loop

# create a vector of times:
#time_vec <- seq(from = 0, to = 35, by = 0.1)

#pop_35_vec <- vector(mode = "numeric", length = length(time_vec))

#for(i in 1:length(time_vec)) {
 # population <- logistic_growth(N0 = 100, K = 6000, r = 0.27, time = time_vec[i])
  #pop_35_vec[i] <- population
#}
```

### Logistic growth equation over several growth rate values

Modeling a population over time and varying growth rate
```{r}
#logistic growth function
logistic_growth <- function(N0, K, r, time) {
  Nt <- K / (1 + ((K - N0) / N0) * exp(-r * time))
  print(Nt)
}

#Create a sequence of growth rate values:
r_seq <- seq(from = 0.20, to = 0.40, by = 0.01)
time_vec <- seq(from = 0, to = 35, by = 0.1)
#Create a matrix to store the outputs
out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

# Now a nested loop
for(i in 1:length(r_seq)) { #outer loop is growth rates
  for(j in 1:length(time_vec)) { #inter loop is time seq values
  pop <- logistic_growth(N0 = 100, K = 6000, r = r_seq[i], time = time_vec[j])
  out_matrix[j,i] <-pop #store the value in the appropriate row and column
    }
}
```

Second Version
```{r}
#short cut cmd + option + x
my_function <- function(oranges, rainbow) {
  (oranges*rainbow/2.4)^2
}
```

Logistic growth computation follow lecture
```{r}
# Create function
logistic_growth <- function(K, N0, r, time) {
  K / (1+ ((K - N0) / N0) * exp(-r * time))
}

#test function:
logistic_growth(K = 5000, N0 = 100, r = 0.07, time = 42)
```

```{r}
#add time vector
time_vec <- seq(from = 0, to = 35, by = 0.1)

# output of estimated population
pop_35 <- logistic_growth(N0 = 40, K = 8000, r = 0.1, time = time_vec)
```

```{r}
# combine data into data frame
pop_time35_df <- data.frame(time_vec, pop_35)
```

data-to-viz.com helps decide what type of of graph is best to visualize data
```{r}
# Now make graph with data frame created 
ggplot( data = pop_time35_df, aes( x = time_vec, y = pop_35)) + geom_line()
```


Logistic growth rate in a loop
```{r}
prop_35_vec <- vector(mode = "numeric", length = length(time_vec))

for(i in seq_along(time_vec)) {
  population <- logistic_growth(N0 = 100, K = 6000, r = 0.27, time = time_vec[i])
  prop_35_vec[i] <- population
}

#view prop_35_vec
```

Now to iterate over multiple growth rates
```{r}
r_seq <- seq(from = 0.20, to = 0.4, by = 0.01)
```

```{r}
#Nested forloop
out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

for(i in seq_along(r_seq)) {
  for( j in seq_along(time_vec)) {
    pop <- logistic_growth(N0 = 100, K = 4000, r = r_seq[i], time = time_vec[j])
    out_matrix[j,i] <- pop
  }
}

### Now some wrangling 
logistic_df <- data.frame(out_matrix, time = time_vec)

# unamed coloumns 
#we want each column to be named growth rate and r_seq
colnames(logistic_df) <- c(paste0("gr_", r_seq), "time")

```

Data is not in tidy data structure single variable for growth rate that is split up into multiples columns 
-  long version of output data frame, for each variable to have a single column
-  we have variables of time, population, and growth rate 

```{r}
logistic_df_long <- logistic_df |> 
  pivot_longer(cols = -time, values_to = "population", names_to = "growth_rate")

#plot
ggplot(data = logistic_df_long, aes(x = time, y = population)) + 
  geom_line(aes(color = growth_rate)) 
```

### A function with warning and an error message 
code runs but still gives warning 

Use `warning` to add a warning.
Use `stop\` to add an error.

Power of watts hitting a tirbine
$$
P=0.3\rho A v^3
$$
rho: air density (kg/m^3)
A: area covered by turbine
v: wind velocity (m/s)

```{r}
calc_windpower <- function(rho, radius, windspeed) {
  
  if(windspeed > 130) {
    warning("Whoa, that's really fast. Are you sure?")
  }
  if(rho > 1.225) {
    warning("That air density is suspicious.")
  }
  if(radius < 0) {
    stop("Rotor radius must be a postive value in (meters).")
  }
  
  0.3*rho*pi*(radius^2)*(windspeed^3)
}

calc_windpower(rho = 1.225, radius = 50, windspeed = 35)

calc_windpower(rho = 1.225, radius = -6, windspeed = 15)

calc_windpower(rho = 3.2, radius = 10, windspeed = 15)



```










